### 2012-09-17 paper.tex
minor rewordings

Tennis is well suited to in-play trading on exchanges since points are played at a steady rate, are clearly separated, and are consistently won and lost by both participants, leading to frequent (and often dramatic) changes in fortune for players but at relatively predictable intervals.  Consequently, the odds can very quickly swing back and forth as traders react to on-court events, generating potential money-making opportunities.  This, in combination with the fact that tennis markets have the ability to offer only a few outcomes (e.g. in a Match Odds market, there are only two outcomes, either one player wins or the other does), ensures its popularity.  Around 80\% of money waged on tennis matches is bet while the match is in progress\footnote{http://www.sportspromedia.com/guest\_blog/peter\_webb\_why\_tennis\_is\_big\_business\_for\_bookmakers}.

@@

Compare with Figure \ref{murrayberrer} which displays the evolution of implied match-winning probabilities for Andy Murray when he played Michael Berrer in the French Open 2011 Men's Third Round.  In particular, observe where the Match Odds probability of Murray winning the match suddenly drops \textit{almost 60\%}.  This phenomenon in the odds data occurred during a \textit{single point in the match}.  Although there are a few other anomalous events could have caused this huge and rapid swing in the market such as a disqualification, the cause could only have been the \textit{injury} that was suffered by Andy Murray.  We quote from the \textit{BBC Sport}\footnote{http://www.bbc.co.uk/sport} live text commentary of the match during this point:

@@

In this case, the injury did not cause Murray to retire from the match and he went on to win in straight sets.  Nevertheless, it is clear the market reacted to this event and its opinion on Murray's chances of winning the match was severely affected.

@@

There have been many past works on the topic of modelling tennis.  \cite{omalley} presents what are considered the \textit{tennis formulae}.  The tennis formulae are a hierarchical series of equations that compute the probability a given player will win a tennis match given the probabilities that the given player and his/her opponent will win any of their service points.  Combined together are individual formulae for the probabilities of winning games, sets, and tiebreaks.  
\cite{servefirst} more comprehensively explore the use of recurrence relations to model tennis, utilising them to calculate probabilities of winning tournaments and also proving explicitly that the probability of winning a set or match does not depend on which player serves first.  \cite{excel} experiment with the same idea in Microsoft Excel.  They investigate using six parameters rather than just the two point-winning probabilities taking into account service faults.  Consequently, for each player they input the probability of a successful first serve, the probability of winning a point on first serve, and the probability of winning a point on second serve.  \cite{willyk} also designed a low-level model of tennis points, taking into account such events as service faults and service aces, in order to predict the outcome of matches.  They tailor their input probabilities to the specific returning capabilities of the given opponent (not an average player), estimated using historical statistics.  They find that their model was able to successfully predict 67\% of outcomes from a sample of 1839 ATP matches played during 2011.  Since we are already adding further complexity with the probability of retirement, we shall concentrate on extending a simpler, two parameter version of a tennis model.  \cite{momentum} continued with an investigation into player momentum in tennis matches by slightly perturbing point-winning probability depending on how much the given player is leading or trailing the match by (essentially introducing a dependency between points).

@@

The granularity of the model is point-level as this is compatible with the structure of standard tennis models.  When a tennis player plays a point, he or she potentially puts great strain on their body.  This strain naturally leads to a chance of an injury occurring.  For the vast majority of points played, the strain is perfectly manageable and does not lead to injury.  For example, Grand Slam winners will hit over 1000 serves during the course of the tournament without issue.  Occasionally however, the body fails to cope with the strain, or the strain is for some reason much more acute than normal (e.g.\ remember Andy Murray twisting his ankle), and an injury occurs.  When a player does get injured during a match, it does not always end in retirement.  Players often soldier on at least for a few points and may even recover from the injury as the match progresses.  
Using this intuition, we present an elegant model for the probability a player will retire \textit{on a given point} in a tennis match:

@@

where $r_n$ is the given player's risk of retirement on point $n$ of the match, $0 \leq \rho \leq 1$, and $X$ and $Y$ are random variables.  Do not confuse $r_n$ with what we are hoping to eventually calculate which is $R_n$, the risk of retiring at some point during the remainder of the match from point $n$.  We say that players start a match with zero probability of retiring ($r_0 = 0$), although in principle it is possible players might start a match with a niggling injury, in which case the magnitude could be inferred from observations of two or more match markets with differing retirement policies.  When setting $r_{n+1}$, we make sure to take the minimum of the calculated $r_{n+1}$ and 1, since retirement risk is a probability and cannot be greater than 1.  The Bernoulli random variable, $X$, with success parameter $c$, models the chance of an injury occurring on a given point.  The random variable, $Y$, models the magnitude of an injury should it occur and is truncated exponentially distributed with rate parameter $\lambda$ and upper bound 1, i.e. minor bumps and bruises are more common than serious, match-ending injuries.  The decay parameter, $\rho$, models the idea that players recover from injuries as matches progress.  For the purposes of avoiding too complex a model, we make the simplifying assumption that $\rho$ and the distributions of $X$ and $Y$ are the same for both players.

@@

A given point in a simulated match has four possible outcomes.  Either player can win the point or either player can retire from the match.  To resolve a point, we generate a random number between 0 and 1 and observe which of the bins (as shown in Figure \ref{bins}), the number falls in.   The match ends if it falls in one of the retirement bins (which is why $r_A$ and $r_B$ are usually 0 else you will rarely be able to get through even a single match without retiring), otherwise one of the players wins the point and the match continues (unless it was match point).

@@

\begin{figure}[H]
\begin{center}
\renewcommand{\arraystretch}{1.1}
\begin{tabular}{|c|c|C{1.5cm}|C{7.3cm}|}
	\hline
	\textbf{Variable} & \textbf{Known} & \textbf{Input / Output} & \textbf{Description} \\ \hline
	$P_A$ & Yes & Input & The probability Player A wins a point on serve (assumed for the moment) \\ \hline
	$P_B$ & Yes & Input & The probability Player B wins a point on serve (assumed for the moment) \\ \hline
	$W_A$ & Yes & Input & The probability Player A wins the match using the standard base tennis model \\ \hline
	$W_B$ & Yes & Input & The probability Player B wins the match using the standard base tennis model	 \\ \hline
	$G_A$ & Yes & Input & The Betfair Set Betting implied probability minus the Betfair Match Odds implied probability for Player A for $G_A \geq 0$ \\ \hline
	$G_B$ & Yes & Input & The Betfair Set Betting implied probability minus the Betfair Match Odds implied probability for Player B for $G_B \geq 0$	 \\ \hline
	- & Yes & Input & The current score in the match \\ \hline \hline
	$W'_A$ & No & Output & The probability Player A wins the match normally given the possibility of retirement \\ \hline
	$W'_B$ & No & Output & The probability Player B wins the match normally given the possibility of retirement \\ \hline
	$R_A$ & No & Output & The probability Player A retires at some point during the remainder of the match (can be categorised by set) \\ \hline
	$R_B$ & No & Output & The probability Player B retires at some point during the remainder of the match (can be categorised by set) \\ \hline \hline
	$c$ & No & Input & The Bernoulli success probability parameter for the truncated hyper-exponential distribution representing the chance a player suffers an injury on any given point	\\ \hline
	$\lambda$ & No & Input & The rate parameter for the truncated hyper-exponential distribution dictating the magnitude of the injury suffered by a player should such an event occur. \\ \hline
	$\rho$ & No & Input & The decay constant representing recovery from injuries in our retirement risk equation	 \\ \hline
\end{tabular}
\renewcommand{\arraystretch}{1}
\end{center}
\caption{Table describing the variables used in our system}
\label{variabletable}
\end{figure}

@@

In order to assess the accuracy of our model, we try it out in a totally artificial environment where we control all the variables.  We ran a single simulated match many times with our chosen set of parameters.  We assume that $P_A$ and $P_B$ remain unchanged throughout the match.  We are looking for 'ideal' scenarios, e.g. where Player A receives an injury during the match but does not retire and still goes on to achieve victory.  When we generate such a match, we recorded into a CSV file the score, the server, and the point-level retirement risks, $r_A$ and $r_B$, at each point during the match.  We then read this information back in, calculating match-winning probabilities at each point in the match.

Figure \ref{artificialwin} shows the effect on simulated markets with varying retirement payout policies of Player A suffering an injury in the second set but still managing to go on to win the match.  We see a sharp drop in the red and orange lines (Match Odds markets with \textit{after one set} and \textit{after one ball} payout policies, respectively) when the injury occurs.  This is followed by recovery, similar to the Murray vs. Berrer example.  The blue line (Match Odds market with \textit{match-completed} policy) ignores retirement risk and therefore does not react to the injury.  Figure \ref{artificialwinrisk} shows the evolution of both the point-level (magenta line) and match-level (green line) retirement risks for Player A in this match.  They are clearly inversely related to the behaviour of the markets; where match-winning probability falls, retirement risk rises, and where match-winning probability recovers, retirement risk decays.

Figure \ref{artificialfirstset} shows a similar situation but where the injury occurs in the first set.  This time we can see the orange line approach the red as we near the end of the first set.  The models the idea that if an injury occurs in the first set of a match, traders in such a market will display increased reluctance to back the player in question as it becomes more likely he or she will finish the set (in case they decide to default afterwards and payouts happen).

Figure \ref{artificialretirement} shows a match where Player A decided not to continue.  Player A does attempt to play one or two more points after the injury and the market anticipates recovery but those hopes are soon dashed.  Note that any discrepancies you might see between the three markets are due to the fact that the modified simulator provides only approximations whereas the base recursive model provides exact solutions (particularly noticeable at deuce or 6-6 in a tiebreak).

@@

Our model appears to produce reasonable results from the beginning of a match given our assumption that players begin matches with no risk of retiring ($r_0^A = r_0^B = 0$).  Our model must also be applicable for any given match state, including the situation where an injury has recently occurred but the affected player is continuing to play.  At this stage, we know the gap in the odds for each player ($G_A$ and $G_B$), but how can we go \textit{backwards} from this and calculate the retirement risk for the \textit{current point}, $r_n^A$ and $r_n^B$ (which will naturally be a lot smaller)?  Well once again, we can use the Nelder-Mead method to approximate these values for each point in the match.  Given this information, our simulator will be able to calculate the match-level retirement risk for each player at each point ($R_n^A$ and $R_n^B$).

@@

Figure \ref{murrayberrerrisk} shows the evolution of Andy Murray's risk of retirement throughout the match.  The green line follows the probability of retirement during the remainder of the match from the given point ($R_n$), whereas the magenta line gives the probability of retiring on a particular point itself ($r_n$).  Our model predicts a peak match-level retirement risk of 45\% during the injury period.  This coincides with a peak point-level retirement risk of 3\%.  Despite the severity of the injury, Murray's position of dominance in the match may be one explanation for why his risk of retirement is not greater. Berrer is still not expected to win the match regardless of his opponent's misfortune.

@@

As the match progresses and Nadal falls further behind, we see spikes in his retirement risk despite there being no large gaps between the Betfair markets.  This happens when the Match Odds market gives a player little chance of winning the match.  We illustrate how this occurs with an example.  Say that on a particular point, the Set Betting market tells us that the implied probability of Nadal winning the match is 0.5.  This means that we can assign $P_A = P_B = 0.645$.  We also happen to know that the difference between the Betfair Set Betting and Match Odds for Nadal is 0.4.  We now look for a value of $r_A$ (retirement risk of Nadal on this point) such that $W_A^{''}$ (probability of winning the match normally with retirement risk after the first set only) is only 0.1.  Since we have $P_A = P_B$ and therefore both players have an equal chance of winning the match if it does not end in retirement, we have $W_B^{''} = 0.1$ and so the sum of the probabilities of either player winning the match normally is only 0.2.  Making the assumption that $r_n^B$ is negligible, we must have that $R_A \approx 0.8$ (retirement risk of Nadal in the match).  This can happen in any situation where $W^{''}$ is small and there is a risk of retirement, like towards the end of this quarter final.

@@

Nonetheless, this match is still a good example of how unpredictable injury occurrence is.  Azarenka's match-level retirement risk still stays relatively low until the injury occurs on point 75.  The risk shoots up to 52\%, peaking at 90\% at retirement on point 93 after a brief dalliance with the possibility of recovery.

@@

We have created a new model for retirement risk in tennis matches and implemented it into a tennis match simulator by incorporating extra parameters approximated using real-world averages and betting odds data as well as additional match outcomes compared to standard tennis models.  To the best of our knowledge, this is the world's first published attempt to create such a model.  Testing the model in a totally artificial environment, we found it was able to mimic the expected patterns of in-play betting markets with different retirement payout policies for matches with varying injury scenarios.

@@

We applied our model to a number of real-world matches (from the point of view of an injured player) using Betfair odds data and produced imitations of the in-play betting markets following different player retirement payout rules.  We find that we can mimic to a good degree of accuracy the progress of the Betfair Set Betting and Match Odds in-play markets throughout matches, i.e. a \textit{match-completed} market and an \textit{after first set} market, respectively.  We attempt to use the Betfair odds data to predict the evolution of an \textit{after one ball} market.  We find that such a market generally correctly produces slightly higher retirement risks than the Betfair Match Odds market when the given match is still in the first set, although it can be somewhat erratic.  The retirement risk values are very sensitive to any gap between the Betfair markets due to the lack of control our system has over retirement risk \textit{after} the first set when the match is still in the first set.  Furthermore, we notice that the system generates retirement risk spikes when the player in question has a low match-winning probability.  The problem underlying these fragilities in our predictions is that the Betfair in-play tennis betting markets are not perfect.  Since this odds data is vital input for our system, it is no surprise that fluctuating, anomalous, and sparse data heavily influences the output of our system.

>>>>

Tennis is well suited to in-play trading on exchanges since points are played at a steady rate, are clearly separated, and are consistently won and lost by both participants, leading to frequent (and often dramatic) changes in fortune for players but at relatively predictable intervals.  Consequently, the odds can very quickly swing back and forth as traders react to on-court events, generating potential money-making opportunities.  This, in combination with the fact that tennis markets have the ability to offer only a few outcomes (e.g.\ in a Match Odds market, there are only two outcomes, either one player wins or the other does), ensures its popularity.  Around 80\% of money waged on tennis matches is bet while the match is in progress\footnote{http://www.sportspromedia.com/guest\_blog/peter\_webb\_why\_tennis\_is\_big\_business\_for\_bookmakers}.

@@

Compare with Figure \ref{murrayberrer} which displays the evolution of implied match-winning probabilities for Andy Murray when he played Michael Berrer in the French Open 2011 Men's Third Round.  In particular, observe where the Match Odds probability of Murray winning the match suddenly drops \textit{almost 60\%}.  This phenomenon in the odds data occurred during a \textit{single point in the match}.  Although there are a few anomalous events which could have led to this huge and rapid swing in the market such as a disqualification, the evidence here weighs heavily towards the \textit{injury} that was suffered by Andy Murray.  We quote from the \textit{BBC Sport}\footnote{http://www.bbc.co.uk/sport} live text commentary of the match during this point:

@@

In this case, the injury did not cause Murray to retire from the match and he went on to win in straight sets, as reflected in the subsequent recovery of his odds to win.  Nevertheless, it is clear the market reacted to this event and its opinion on Murray's chances of winning the match was severely affected.

@@

There have been many past works on the topic of modelling tennis.  \cite{omalley} presents what are considered the \textit{tennis formulae}.  The tennis formulae are a hierarchical series of equations that compute the probability a given player will win a tennis match given the probabilities that the given player and his/her opponent will win any of their service points.  Combined together are individual formulae for the probabilities of winning games, sets, and tiebreaks.

\cite{servefirst} more comprehensively explore the use of recurrence relations to model tennis, utilising them to calculate probabilities of winning tournaments and also proving explicitly that the probability of winning a set or match does not depend on which player serves first.  \cite{excel} experiment with the same idea in Microsoft Excel.  They investigate using six parameters rather than just the two point-winning probabilities taking into account service faults.  Consequently, for each player they input the probability of a successful first serve, the probability of winning a point on first serve, and the probability of winning a point on second serve.  \cite{willyk} also designed a low-level model of tennis points, taking into account such events as service faults and service aces, in order to predict the outcome of matches.  They tailor their input probabilities to the specific returning capabilities of the given opponent (not an average player), estimated using historical statistics.  They find that their model was able to successfully predict 67\% of outcomes from a sample of 1839 ATP matches played during 2011.  Since we are already adding further complexity with the probability of retirement, we shall concentrate on extending a simpler, two parameter version of a tennis model.  \cite{momentum} continued with an investigation into player momentum in tennis matches by slightly perturbing point-winning probability depending on how much the given player is leading or trailing the match by (essentially introducing a dependency between points).

@@

The granularity of the model is point-level as this is compatible with the structure of standard tennis models.  When a tennis player plays a point, he or she potentially puts great strain on their body.  This strain naturally leads to a chance of an injury occurring.  For the vast majority of points played, the strain is perfectly manageable and does not lead to injury.  For example, Grand Slam winners can hit over 1000 serves during the course of the tournament without issue.  Occasionally however, the body fails to cope with the strain, or the strain is for some reason much more acute than normal (e.g.\ remember Andy Murray twisting his ankle), and an injury occurs.  When a player does get injured during a match, it does not always end in retirement.  Players often soldier on at least for a few points and may even recover from the injury as the match progresses.  
Using this intuition, we present an elegant model for the probability a player will retire \textit{on a given point} in a tennis match:

@@

where $r_n$ is the given player's risk of retirement on point $n$ of the match, $0 \leq \rho \leq 1$, and $X$ and $Y$ are random variables.  Do not confuse $r_n$ with what we are hoping to eventually calculate which is $R_n$, the risk of retiring at some point during the remainder of the match from point $n$.  We say that players start a match with zero probability of retiring ($r_0 = 0$), although in principle it is possible players might start a match with a niggling injury, in which case the magnitude could be inferred from observations of two or more match markets with differing retirement policies.  When setting $r_{n+1}$, we make sure to take the minimum of the calculated $r_{n+1}$ and 1, since retirement risk is a probability and cannot be greater than 1.  The Bernoulli random variable, $X$, with success parameter $c$, models the chance of an injury occurring on a given point.  The random variable, $Y$, models the magnitude of an injury should it occur and is truncated exponentially distributed with rate parameter $\lambda$ and upper bound 1, i.e. minor bumps and bruises are more common than serious, match-ending injuries.  The decay parameter, $\rho$, models the idea that players recover from injuries as matches progress.  For the purpose of avoiding too complex a model, we make the simplifying assumption that $\rho$ and the distributions of $X$ and $Y$ are the same for both players.

@@

A given point in a simulated match has four possible outcomes.  Either player can win the point or either player can retire from the match.  To resolve a point, we generate a random number between 0 and 1 and observe which of the bins (as shown in Figure \ref{bins}) the number falls in.  The match ends if it falls in one of the retirement bins (which is why $r_A$ and $r_B$ are usually 0 else you will rarely be able to get through even a single match without retiring), otherwise one of the players wins the point ($P_A$ is the point-winning probability on serve of Player A) and the match continues (unless it was their match point).

@@

\begin{figure}[H]
\begin{center}
\renewcommand{\arraystretch}{1.1}
\begin{tabular}{|c|c|C{1.5cm}|C{7.3cm}|}
	\hline
	\textbf{Variable} & \textbf{Known} & \textbf{Input / Output} & \textbf{Description} \\ \hline
	$P_A$ & Yes & Input & The probability Player A wins a point on serve (assumed for the moment) \\ \hline
	$P_B$ & Yes & Input & The probability Player B wins a point on serve (assumed for the moment) \\ \hline
	$W_A$ & Yes & Input & The probability Player A wins the match using a standard tennis model \\ \hline
	$W_B$ & Yes & Input & The probability Player B wins the match using a standard tennis model	 \\ \hline
	$G_A$ & Yes & Input & The Betfair Set Betting implied probability minus the Betfair Match Odds implied probability for Player A for $G_A \geq 0$ \\ \hline
	$G_B$ & Yes & Input & The Betfair Set Betting implied probability minus the Betfair Match Odds implied probability for Player B for $G_B \geq 0$	 \\ \hline
	- & Yes & Input & The current score in the match \\ \hline \hline
	$W'_A$ & No & Output & The probability Player A wins the match normally given the possibility of retirement \\ \hline
	$W'_B$ & No & Output & The probability Player B wins the match normally given the possibility of retirement \\ \hline
	$R_A$ & No & Output & The probability Player A retires at some point during the remainder of the match (can be categorised by set) \\ \hline
	$R_B$ & No & Output & The probability Player B retires at some point during the remainder of the match (can be categorised by set) \\ \hline \hline
	$c$ & No & Input & The Bernoulli success probability parameter for the truncated hyper-exponential distribution representing the chance a player suffers an injury on any given point	\\ \hline
	$\lambda$ & No & Input & The rate parameter for the truncated hyper-exponential distribution dictating the magnitude of the injury suffered by a player should such an event occur. \\ \hline
	$\rho$ & No & Input & The decay constant representing recovery from injuries in our retirement risk equation	 \\ \hline
\end{tabular}
\renewcommand{\arraystretch}{1}
\end{center}
\caption{Table describing the variables used in our system}
\label{variabletable}
\end{figure}

@@

In order to assess the accuracy of our model, we try it out in a totally artificial environment where we control all the variables.  We run a single simulated match many times with our chosen set of parameters.  We assume that $P_A$ and $P_B$ remain unchanged throughout the match.  We are looking for `ideal' scenarios, e.g. where Player A receives an injury during the match but does not retire and still goes on to achieve victory.  When we generate such a match, we record into a CSV file the score, the server, and the point-level retirement risks, $r_A$ and $r_B$, at each point during the match.  We then read this information back in, calculating match-winning probabilities at each point in the match.

Figure \ref{artificialwin} shows the effect on simulated markets with varying retirement payout policies of Player A suffering an injury in the second set but still managing to go on to win the match.  We see a sharp drop in the red and orange lines (Match Odds markets with \textit{after one set} and \textit{after one ball} payout policies, respectively) when the injury occurs.  Traders now fear a retirement and are less willing to back Player A to win the match.  This is followed by recovery, similar to the Murray vs. Berrer example, as traders' faith in a Player A victory is slowly restored.  The blue line (Match Odds market with \textit{match-completed} policy) ignores retirement risk and therefore does not react to the injury.  Figure \ref{artificialwinrisk} shows the evolution of both the point-level (magenta line) and match-level (green line) retirement risks for Player A in this match.  They are clearly inversely related to the behaviour of the markets; where match-winning probability falls, retirement risk rises, and where match-winning probability recovers, retirement risk decays.

Figure \ref{artificialfirstset} shows a similar situation but where the injury occurs in the first set.  This time the red line does not fall as drastically as the orange and the two markets merge as we approach the end of the first set.  This models the idea that if an injury occurs in the first set of a match, traders in an \textit{after first set} market will only display increased reluctance to back the player in question as it becomes more likely he or she will finish the set (in case they decide to default afterwards and payouts happen).

Figure \ref{artificialretirement} shows a match where Player A decided not to continue.  Player A does attempt to play one or two more points after the injury and the market anticipates recovery but those hopes are soon dashed.  Note that any discrepancies you might see between the three markets are due to the fact that the modified simulator provides only approximations whereas the standard model provides exact solutions (particularly noticeable at deuce or 6-6 in a tiebreak).

@@

Our model appears to produce reasonable results from the beginning of a match given our assumption that players begin matches with no risk of retiring ($r_0^A = r_0^B = 0$).  Our model must also be applicable for any given match state, including the situation where an injury has recently occurred but the affected player is continuing to play.  At this stage, we know the gap in the odds for each player ($G_A$ and $G_B$), but how can we go \textit{backwards} from this and calculate the retirement risk for the \textit{current point}, $r_n^A$ and $r_n^B$ (which will naturally be a lot smaller)?  Once again, we can use the Nelder-Mead method to approximate these values for each point in the match.  Given this information, our simulator will be able to calculate the match-level retirement risk for each player at each point ($R_n^A$ and $R_n^B$).

@@

Figure \ref{murrayberrerrisk} shows the evolution of Andy Murray's risk of retirement throughout the match.  The green line follows the probability of retirement during the remainder of the match from the given point ($R_n$), whereas the magenta line gives the probability of retiring on a particular point itself ($r_n$).  Our model predicts a peak match-level retirement risk of 45\% during the injury period.  This coincides with a peak point-level retirement risk of 3\%.  Despite the severity of the injury, Murray's position of dominance in the match may be one explanation for why his risk of retirement is not greater. Berrer is still not expected to win the match irrespective of his opponent's misfortune.

@@

As the match progresses and Nadal falls further behind, we see spikes in his retirement risk despite there being no large gaps between the Betfair markets.  This happens when the Match Odds market gives a player little chance of winning the match.  We illustrate how this occurs with an example.  Say that on a particular point, the Set Betting market tells us that the implied probability of Nadal winning the match is 0.5.  This means that we can assign $P_A = P_B = 0.645$.  We also happen to know that the difference between the Betfair Set Betting and Match Odds markets for Nadal is 0.4.  We now look for a value of $r_A$ (retirement risk of Nadal on this point) such that $W_A^{''}$ (probability of winning the match normally with retirement risk after the first set only) is only 0.1.  Since we have $P_A = P_B$ and therefore both players have an equal chance of winning the match if it does not end in retirement, we have $W_B^{''} = 0.1$ and so the sum of the probabilities of either player winning the match normally is only 0.2.  Making the assumption that $r_n^B$ is negligible, we must have that $R_A \approx 0.8$ (retirement risk of Nadal in the match).  This can happen in any situation where $W^{''}$ is small and there is a risk of retirement, like towards the end of this quarter final.

@@

Nonetheless, this match is still a good example of how unpredictable injury occurrence is.  Azarenka's match-level retirement risk still stays relatively low until the injury occurs on point 75.  The risk shoots up to 54\%, peaking at 89\% at retirement on point 93 after a brief dalliance with the possibility of recovery.

@@

We have created a new model for retirement risk in tennis matches and implemented it into a tennis match simulator by incorporating extra parameters approximated using real-world averages and betting odds data as well as additional match outcomes compared to standard tennis models.  To the best of our knowledge, this is the world's first published attempt to create such a model.  Testing the model in a completely artificial environment, we found it was able to mimic the expected patterns of in-play betting markets with different retirement payout policies for matches with varying injury scenarios.

@@

We applied our model to a number of real-world matches (from the point of view of an injured player) using Betfair odds data and produced imitations of the in-play betting markets following different player retirement payout rules.  We find that we can mimic to a good degree of accuracy the progress of the Betfair Set Betting and Match Odds in-play markets throughout matches, i.e. a \textit{match-completed} market and an \textit{after first set} market, respectively.  We attempt to use the Betfair odds data to predict the evolution of an \textit{after one ball} market.  We find that such a market generally correctly produces slightly lower match-winning implied probabilities than the Betfair Match Odds market when the given match is still in the first set, although it can be somewhat erratic.  The retirement risk values are very sensitive to any gap between the Betfair markets due to the lack of control our system has over retirement risk \textit{after} the first set when the match is still in the first set.  Furthermore, we notice that the system generates retirement risk spikes when the player in question has a low match-winning probability.  The problem underlying these fragilities in our predictions is that the Betfair in-play tennis betting markets are not perfect.  Since this odds data is vital input for our system, it is no surprise that fluctuating, anomalous, and sparse data heavily influences the output of our system.
